---
title: "Functions"
output: html_document
---

```{r setup}
library(tidyverse)

df <- tibble::tibble(
a = c(rnorm(9), -99),
b = c(-999, -99, rnorm(8)),
c = c(0, rnorm(9)),
d = rnorm(10)
)
(df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a)))
(df$b <- (df$b - min(df$b)) / (max(df$b) - min(df$b)))
(df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c)))
(df$d <- (df$d - min(df$d)) / (max(df$d) - min(df$c)))
```

Write a function to automate the calculation of the rescale:
```{r}
#	Rescale	to	[0,	1]
0	+	(1	-	0)	*	((df$a	-	min(df$a))	/	(max(df$a)	-	min(df$a)))

#	Rescale	to	[-1,	1]
-1	+	(1	-	-1)	*	((df$b	-	min(df$b))	/	(max(df$b)	-	min(df$b)))

#	Rescale	to	[0,	10]
0	+	(10	-	0)	*	((df$c	-	min(df$c))	/	(max(df$c)	-	min(df$c)))

rescale<- function(x, min, max) {
  rng<- range(x, na.rm=TRUE, finite=TRUE)
  min + (max- min) *((x-rng[1])/(rng[2]-rng[1]))
}

rescale <- function(x, min, max) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  min + (max - min) * ((x - rng[1]) / (rng[2] - rng[1]))
}

```


##Your Turn 1

Rerun the tibble to revert back to the original data.  Write a function to replace -99 with NA, called `fix_missing()`

```{r}
df <- tibble::tibble(
a = c(rnorm(9), -99),
b = c(-999, -99, rnorm(8)),
c = c(0, rnorm(9)),
d = rnorm(10)
)


#Here are the repeated calculations
df$a[df$a==-99]<-NA
df$b[df$b==-99]<-NA
df$c[df$c==-99]<-NA
df$d[df$d==-99]<-NA



```


Now let's write a function that allows us to easily adapt our rescaling to include a range (min, max) to modify the data.
```{r}
#Rescale to [0, 1]
0 + (1 - 0) * ((df$a - min(df$a)) / (max(df$a) - min(df$a)))

# Rescale to [-1, 1]
-1 + (1 - -1) * ((df$b - min(df$b)) / (max(df$b) - min(df$b)))

# Rescale to [0, 10]
0 + (10 - 0) * ((df$c - min(df$c)) / (max(df$c) - min(df$c)))

```


## Your Turn 2
Expand your function from 'Your Turn 1' to allow for any possible missing value type.
```{r}
df$a[df$a==-99]<-NA
df$b[df$b==-99]<-NA
df$c[df$c==-99]<-NA
df$d[df$d==-99]<-NA

fix_missing <- function(x, missing_val=c(-99,-999, -0)) {
  x[x %in% missing_val ]<-NA
  x
}

fix_missing(df$a)
fix_missing(df$b, missing_val=-999)
fix_missing(df$c, missing_val= 0)
```



# Take Aways

To write a function, 

1. Write code that solves the problem for a real object  
2. Wrap the code in `function(){}` to save it  
3. Add the name of the real object as the function argument 
4. Write code in a way that is future-oriented.
5. Be consisentent.

This sequence will help prevent bugs in your code (and reduce the time you spend correcting bugs). 

```{r}
# Rescale to [0, 1]  
rescale(df$a)

# Rescale to [-1, 1]  
rescale(df$b, min = -1, max = 1)

# Rescale to [0, 10]  
rescale(df$c, max = 10)

```

